---
title: "R örnek 1"
author: "Ümit Mert Çağlar"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_notebook: default
---


# Step 1: Install and Load Necessary Packages


```{r eval=FALSE, include=FALSE}
# Install necessary packages (if not already installed)
install.packages("tidyverse")
install.packages("ggplot2")
install.packages("dplyr")
install.packages("readr")
```


Gerekli kütüphaneler yüklenir.


```{r}
# Load the libraries
library(tidyverse)
library(ggplot2)
library(dplyr)
library(readr)
```


# Step 2: Load the Data


```{r}
# Load a dataset (for example, the built-in 'mtcars' dataset)
data(mtcars)

# Alternatively, you can load your own dataset from a CSV file
# mtcars <- read_csv("path_to_your_file.csv")

# View the first few rows of the dataset
head(mtcars)
```

# Step 3: Basic Data Exploration

```{r}
# Get a summary of the dataset
summary(mtcars)

# Check the structure of the dataset
str(mtcars)

# Check for missing values
sum(is.na(mtcars))

# Get the dimensions of the dataset (rows x columns)
dim(mtcars)

# Get the column names
colnames(mtcars)
```


# Step 4: Data Cleaning (if necessary)

```{r}
# Example: Replace missing values with the mean (if any)
mtcars <- mtcars %>%
  mutate(across(everything(), ~ ifelse(is.na(.), mean(., na.rm = TRUE), .)))

# Check again for missing values
sum(is.na(mtcars))
```



# Step 5: Data Visualization

```{r}
# Scatter plot of mpg vs. hp
ggplot(mtcars, aes(x = hp, y = mpg)) +
  geom_point() +
  labs(title = "Scatter plot of MPG vs. Horsepower",
       x = "Horsepower",
       y = "Miles per Gallon")

# Boxplot of mpg by number of cylinders
ggplot(mtcars, aes(x = as.factor(cyl), y = mpg)) +
  geom_boxplot() +
  labs(title = "Boxplot of MPG by Number of Cylinders",
       x = "Number of Cylinders",
       y = "Miles per Gallon")

# Histogram of mpg
ggplot(mtcars, aes(x = mpg)) +
  geom_histogram(binwidth = 5, fill = "blue", color = "black") +
  labs(title = "Histogram of MPG",
       x = "Miles per Gallon",
       y = "Frequency")
```

# Step 6: Basic Statistical Analysis

```{r}
# Correlation matrix
cor(mtcars)

# Linear regression model: mpg as a function of hp and wt
model <- lm(mpg ~ hp + wt, data = mtcars)
summary(model)

# ANOVA to compare means of mpg across different cylinder groups
anova_result <- aov(mpg ~ as.factor(cyl), data = mtcars)
summary(anova_result)
```


# Step 7: Save the Cleaned Data (Optional)

```{r}
# Save the cleaned dataset to a new CSV file
write_csv(mtcars, "cleaned_mtcars.csv")
```


# Step 8: Generate a Report (Optional)
You can use R Markdown to generate a report that includes your code, visualizations, and analysis.


```{r}
# Install R Markdown if not already installed
# install.packages("rmarkdown")
```


